Uppangala tree functionnal traits: Preliminray Results

### 2016-06-20 inst/doc/PreliminaryResults.Rmd
v0.1.9012 CWT.R, null_model.R, RLQ, SAC, PhyloS, CWV, SEM

@@

```{r Traits, eval=T}
rm(list = ls()); invisible(gc())
library(Uppangala)
PFT <- genPFT()
captions <- genCaptions()
table <- data.frame(
  traits = c('Thickness', 'Leaf area', 'Leaf dry matter content', 'Specific leaf area', 'Wood density'),
  abbreviation = c('Thick', 'LA', 'LDMC', 'SLA', 'WD'),
  range = c(paste(range(PFT$Thick, na.rm = T), collapse = '-'),
            paste(round(range(PFT$LA, na.rm = T)), collapse = '-'),
            paste(round(range(PFT$LDMC, na.rm = T)), collapse = '-'),
            paste(round(range(PFT$SLA, na.rm = T), 2), collapse = '-'),
            paste(round(range(PFT$WD, na.rm = T), 2), collapse = '-')),
  `mean (sd)` = c(paste0(round(mean(PFT$Thick, na.rm = T)), ' (',
                         round(sd(PFT$Thick, na.rm = T)), ')'),
                  paste0(round(mean(PFT$LA, na.rm = T)), ' (',
                         round(sd(PFT$LA, na.rm = T)), ')'),
                  paste0(round(mean(PFT$LDMC, na.rm = T)), ' (',
                         round(sd(PFT$LDMC, na.rm = T)), ')'),
                  paste0(round(mean(PFT$SLA, na.rm = T),2), ' (',
                         round(sd(PFT$SLA, na.rm = T),2), ')'),
                  paste0(round(mean(PFT$WD, na.rm = T),2), ' (',
                         round(sd(PFT$WD, na.rm = T),2), ')')),
  unit = c('um', 'mm2', 'mg.g-1', 'm2.kg-1', 'g.cm-3'),
  strategy = c('Leaf defense vs. investment',
               'Leaf ressource capture vs.investment',
               'Leaf defense vs. investment',
               'Leaf ressource acquisiton vs. defense',
               'Stem transport, structure and defense')
  )
kable(table,  label="traits", row.names = F)
```

@@

```{r CV}
PFT_ind <- aggregate(PFT, by = list(PFT$Tree), FUN = mean, na.rm = T)
row.names(PFT_ind) <- PFT_ind$Group.1
SpT <- unique(PFT[1:2])
Sp <- SpT$SP
names(Sp) <- SpT$Tree
SpCodeT <- unique(PFT[2:3])
SpCode <- SpCodeT$Sp_Code
names(SpCode) <- SpCodeT$SP
CET <- unique(PFT[c(1,4)])
CE <- CET$CE
names(CE) <- CET$Tree
PFT_ind$SP <- Sp[row.names(PFT_ind)]
PFT_ind$Sp_Code <- SpCode[as.character(PFT_ind$SP)]
PFT_ind$CE <- CE[row.names(PFT_ind)]
PFT_ind <- PFT_ind[-c(1:2)]
cv <- data.frame(row.names = c('Thick', 'LA', 'LDMC', 'SLA', 'WD'))
cv$Individual <- unlist(lapply(
  as.list(row.names(cv)),
  function(y){mean(aggregate(PFT[which(names(PFT) ==  y)], by = list(PFT$Tree), function(x){sd(x, na.rm = T) / mean(x, na.rm = T)})[,2], na.rm = T)}))
cv$Species <- unlist(lapply(
  as.list(row.names(cv)),
  function(y){mean(aggregate(PFT_ind[which(names(PFT_ind) ==  y)], by = list(PFT_ind$SP), function(x){sd(x, na.rm = T) / mean(x, na.rm = T)})[,2], na.rm = T)}))
cv$Community <- unlist(
  lapply(as.list(row.names(cv)),
         function(y){sd(PFT[,which(names(PFT) ==  y)], na.rm = T) /
             mean(PFT[,which(names(PFT) ==  y)], na.rm = T)}))
cv <- round(cv, 2)
cv['WD','Individual'] <- ''
kable(cv)
```

@@

```{r PCA PFT}
library(vegan)
pca <- princomp(~ LA + WD + Thick + LDMC + SLA, data = PFT_ind, cor = T)
lambda <- pca$sdev * sqrt(nrow(pca$scores))
pca.var = apply(pca$score, 2, var)
PCAscores <- as.data.frame(pca$scores)[1:2] # Adding PCA axis in PFT values
names(PCAscores) <- c('WES', 'LES')
PCAscores$ID <- row.names(PCAscores)
plot(t(t(pca$scores)/lambda),pch=16, col = 'lightgrey',
     # xlab = paste('Axe 1 -', round(lambda[1]/sum(lambda)*100, 2), '%'),
     # ylab = paste('Axe 2 -', round(lambda[2]/sum(lambda)*100, 2), '%'))
     xlab = paste('Axe 1 -', round(pca.var[1]/sum(pca.var)*100, 2), '%'),
     ylab = paste('Axe 2 -', round(pca.var[2]/sum(pca.var)*100, 2), '%'))
# as.factor(Species[PFT_ind[row.names(pca$scores),2],6])
par(new=T)
Rot <- t(t(pca$loadings)*lambda)
XLIM <- c(-max(abs(Rot[,1])),max(abs(Rot[,1])))
XLIM <- XLIM+(XLIM*0.2)
plot(Rot,col=4,axes=FALSE,xlim=XLIM,ylim=XLIM,pch="",xlab = "", ylab = "")
lines(XLIM, c(0,0), lty = 2, col = 'firebrick', lwd = 2)
lines(c(0,0), XLIM, lty = 2, col = 'green', lwd = 2)
legend('topright', c('Leaf economic spectrum', 'Wood economic spectrum'),
       text.col = c("green", "firebrick"))

@@

```{r env-com}
library(raster)
Env <- genEnv()
Env$Rivers <- readAll(Env$Rivers)
Trees <- genTrees()
Species <- genSpecies()
PFT_sp <- aggregate(PFT_ind, by = list(PFT_ind$SP), FUN = mean, na.rm = T)
row.names(PFT_sp) <- PFT_sp$Group.1
PFT_sp$Sp_Code <- SpCode[row.names(PFT_sp)]
PFT_sp <- PFT_sp[-c(1:2,4)]
Trees$SpCode <- tolower(Trees$SpCode)
Trees$ID <- row.names(Trees)
Trees <- merge(Trees, PFT_sp, by.x = 'SpCode', by.y = 'Sp_Code')
row.names(Trees) <- Trees$ID
Trees <- Trees[-which(names(Trees) == 'ID')]
com <- quadrats()
Trees$com <- NA
XY <- data.frame(X = Trees$X, Y = Trees$Y)
XY <- XY[-which(is.na(XY$X)),] # Removing NA
coordinates(XY) <- ~ X + Y
proj4string(XY) <- crs(com)
Trees$com[which(!is.na(Trees$X))] <- (XY %over% com)[,1]
Trees$com[which(!is.na(Trees$com))] <- paste0('C', Trees$com[which(!is.na(Trees$com))])
com@data$id <- paste0('C', com@data$id)
com@data$DEM <- extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
Deriv <- DEMderiv(raster(com, 'DEM'), c('slope', 'curvature', 'plancurvature', 'profcurvature', 'aspect'))
com@data <- cbind(com@data, extract(Deriv, as(com, "SpatialPolygons"), df = T, fun = mean)[-1])
com@data$Wetness <- extract(DEMderiv(Env$DEM, 'wetness'), as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$SouthWesterness <- sin((com@data$Aspect + 225) / 180 * pi)
com@data$NorthEasterness <- cos((com@data$Aspect + 225) / 180 * pi)
com@data$Rivers <- extract(Env$Rivers, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$topo <- as.factor(with(com@data, topo_classes(Slope, Curvature, Aspect)))
com@data$Canopy <- extract(Env$Canopy, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$BA <- aggregate(Trees$`2013_girth`, list(Trees$com), function(x){sum(x, na.rm = T)/2*pi})[,2]
com@data$StD <- aggregate(rep(1, length(Trees$SpCode)), list(Trees$com), sum)[,2]
rm(Deriv, XY)
```

@@

To study abiotic factors we choose 3 available variables: canopy height (canopy), basal area (BA), and stem density (StD) *(we may add growth in the future)*. Canopy height is issued from a digital canopy model built with a LIDAR campaign on Uppangala plots with a resolution $1*1 m^2$. Both basal area and stem density are calculated with 2013 Uppangala plots girth inventory only on stem with a diameter above 10 centimeters at breast height ($DBH>10$). Consequently basal area end stem density are highly correlated (Spearman's $rho = 0.84$, $p-value < 0.001$). So we removed stem density to avoid multicolinearity, and we kept canopy height that wasn't correlated to basal area (Spearman's $rho = -0.09$, $p-value > 0.1$).

For each species the potential strata was defined. Because species were defined in several strata we kept the lowest strata as the species strata. We tested the strata effect on each trait with an anova combined to a tukey test, but no trait show a significant variation (all F values weren't significanlty different from 0, $p-values>0.1$).

```{r strata effect, eval=FALSE, include=FALSE, cache=F}
AOV <- list(SLA = aov(PFT_sp$SLA ~ Species[PFT_sp$Sp_Code,]$Strata), 
            WD = aov(PFT_sp$WD ~ Species[PFT_sp$Sp_Code,]$Strata),
            LDMC = aov(PFT_sp$LDMC ~ Species[PFT_sp$Sp_Code,]$Strata),
            LA = aov(PFT_sp$LA ~ Species[PFT_sp$Sp_Code,]$Strata),
            Thick = aov(PFT_sp$Thick ~ Species[PFT_sp$Sp_Code,]$Strata))
lapply(AOV, summary)
# lapply(AOV, TukeyHSD)
```

@@

```{r Com}
CWM <- aggregate(Trees, by = list(Trees$com), mean, na.rm = T) #CWM abundance
names(CWM)[1] <- 'id'
CWM <- CWM[which(names(CWM) %in% c('id', 'LES', 'SLA', 'WES', 'LDMC', 'WD', 'Thick', 'LA'))]
com_abund <- merge(com, CWM)
Trees_bin <- Trees[-which(duplicated(cbind(Trees$SpCode, Trees$com))),] #CWM binary
CWM <- aggregate(Trees_bin, by = list(Trees_bin$com), mean, na.rm = T)
names(CWM)[1] <- 'id'
CWM <- CWM[which(names(CWM) %in% c('id', 'LES', 'SLA', 'WES', 'LDMC', 'WD', 'Thick', 'LA'))]
com_bin <- merge(com, CWM)
library(plyr)
TreesBA <- Trees[-which(is.na(Trees$`2013_girth`)),] #CWM ba
CWM <- ddply(TreesBA, .(com), function(x){
  weighted.mean(x$SLA, x$`2013_girth`/2*pi, na.rm = T)})
CWM <- data.frame(id = CWM$com,
                  Thick = CWM$V1,
                  LA = ddply(TreesBA, .(com), function(x){weighted.mean(x$LA, x$`2013_girth`/2*pi, na.rm = T)})[,2],
                  LDMC = ddply(TreesBA, .(com), function(x){weighted.mean(x$LDMC, x$`2013_girth`/2*pi, na.rm = T)})[,2],
                  SLA = ddply(TreesBA, .(com), function(x){weighted.mean(x$SLA, x$`2013_girth`/2*pi, na.rm = T)})[,2],
                  WD = ddply(TreesBA, .(com), function(x){weighted.mean(x$WD, x$`2013_girth`/2*pi, na.rm = T)})[,2],
                  WES = ddply(TreesBA, .(com), function(x){weighted.mean(x$WES, x$`2013_girth`/2*pi, na.rm = T)})[,2],
                  LES = ddply(TreesBA, .(com), function(x){weighted.mean(x$LES, x$`2013_girth`/2*pi, na.rm = T)})[,2])
com_ba <- merge(com, CWM)
```

Among the whole 10 hectares plot from Uppangala, 240 spatial communities where defined with $20*20 m^2$ quadrats. For all those community, community weighted means [CWM, sensu @violle_let_2007] were computed for each functional traits and from the species scores obtained on the 2 first PCA axes representing the leaf and the wood economic spectrum (WES, LES). Weighted mean were realized by abundance, presence-absence, or basal area.

Spatial variograms of community weighted means didn't show a clear plateau (see figure 4) but suggest a spatial organization at a scale of $70 m$.

```{r Variograms}
library(gstat)
dist <- variogram(WD~1, as(com_abund, 'SpatialPointsDataFrame'))$dist
plot(c(1:length(dist)) ~ dist, xlim = c(0, max(dist)), ylim = c(0.5,1), pch = '', xlab = 'distance (m)', ylab = 'relative variance (%)')
com.L <- list(Abundance = com_abund,
              'Presence-absence' = com_bin,
              'Basal area' = com_ba)
for(j in 1:length(com.L)){
  var <- data.frame(lapply(lapply(as.list(c('WD', 'Thick', 'LA', 'LDMC', 'SLA', 'WES', 'LES')), function(x){variogram(formula(paste(x, '~ 1')), as(com.L[[j]], 'SpatialPointsDataFrame'))}), function(x){x[3]}))
  var <- data.frame(apply(var, 2, function(x){x/max(x)}))
  names(var) <- c('WD', 'Thick', 'LA', 'LDMC', 'SLA', 'WES', 'LES')
  colramp <- rainbow(length(var))
  for(i in 1:length(var)){
    lines(var[,i] ~ dist, col = colramp[i], lty = j + 1)
  }
}
legend('bottomright', names(var), fill = colramp)
legend('bottomleft', names(com.L), lty = c(2:4))
```

Figure 4: Trait variograms for all communities.*Variogram was initiated on* $20*20 m^2$ *communities with the wood density weighted mean, and seems to reach a plateau around* $70 m$.

Links between community weighted means and environment, for both abiotic and biotic factors, were studied by linear regression on continuous variables and anova with factors. Linear regression models were chosen by a step mesure (in both direction) on a first regression with selected variables (slope, curvature, wetness and southwesterness for abiotic factors, and canopy height and basal area for biotic one). A null model was built by randomizing traits among species (permuting raw in `Trait*Species` matrix). Deviance of models selected by the step mesure was calculated 999 times and compare to observed deviance. Results for abundance, presence-absence and basal area community weighted mean for all traits are presented in tables 3 and 4 (respectively for abiotic and biotic factors).

@@

```{r topo results}
reg <- '~ Slope + Curvature + Wetness + SouthWesterness'
LM <- lapply(list(com_abund, com_bin, com_ba),
             function(x) {with(x@data, list(LES = lm(formula(paste('LES', reg))),
                                            logSLA = lm(formula(paste('log(SLA)', reg))),
                                            WES = lm(formula(paste('WES', reg))),
                                            WD = lm(formula(paste('WD', reg))),
                                            LDMC = lm(formula(paste('LDMC', reg))),
                                            logLA = lm(formula(paste('log(LA)', reg))),
                                            Thick = lm(formula(paste('Thick', reg)))))})
names(LM) <- c('Abundance', 'Presence/Absence', 'Basal area')
LM <- lapply(LM, function(y){lapply(y, function(x){
  step(x, direction = 'both', trace = 0)})})
TreesL <- list(Trees, Trees_bin, TreesBA)
comL <- list(com_abund, com_bin, com_ba)
baL <- list(F, F, T)
model <- mapply(function(lm, trees, com, ba){
  null_model(formulas = lapply(lm, function(x){as.character(x$call)[2]}),
             ba = ba, test = 'lm deviance', PFT_sp = PFT_sp, Trees = trees,
             com = com, SpCode = SpCode, n = 1000, plot = F, count = F, time = F)},
  LM, TreesL, comL, baL, SIMPLIFY = F)

kable(regression_table(LM, model, reg))
```

Table 3: Step mesure on abiotic factors and null models summary for traits community weighted means. *Linear model results for each trait community weighted means (by abundance, presence-absence and basal area) after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW). Model deviance was compared to 999 null models deviance in which traits were randomized among species.*

@@

>>>>

# *News*
Add from the june 20th version :

- RLQ to investigate individuals filters
- Spatial autocorrelation for CWM and models residuals with Moran's I
- Phylosignal
- CWV
- Structural Equation Modelling (SEM)

@@

```{r 1 Traits}
rm(list = ls()); invisible(gc())
library(Uppangala)
PFT <- genPFT()
captions <- genCaptions()
table <- data.frame(
  traits = c('Thickness', 'Leaf area', 'Leaf dry matter content', 'Specific leaf area', 'Wood density'),
  abbreviation = c('Thick', 'LA', 'LDMC', 'SLA', 'WD'),
  range = c(paste(range(PFT$Thick, na.rm = T), collapse = '-'),
            paste(round(range(PFT$LA, na.rm = T)), collapse = '-'),
            paste(round(range(PFT$LDMC, na.rm = T)), collapse = '-'),
            paste(round(range(PFT$SLA, na.rm = T), 2), collapse = '-'),
            paste(round(range(PFT$WD, na.rm = T), 2), collapse = '-')),
  `mean (sd)` = c(paste0(round(mean(PFT$Thick, na.rm = T)), ' (',
                         round(sd(PFT$Thick, na.rm = T)), ')'),
                  paste0(round(mean(PFT$LA, na.rm = T)), ' (',
                         round(sd(PFT$LA, na.rm = T)), ')'),
                  paste0(round(mean(PFT$LDMC, na.rm = T)), ' (',
                         round(sd(PFT$LDMC, na.rm = T)), ')'),
                  paste0(round(mean(PFT$SLA, na.rm = T),2), ' (',
                         round(sd(PFT$SLA, na.rm = T),2), ')'),
                  paste0(round(mean(PFT$WD, na.rm = T),2), ' (',
                         round(sd(PFT$WD, na.rm = T),2), ')')),
  unit = c('um', 'mm2', 'mg.g-1', 'm2.kg-1', 'g.cm-3'),
  strategy = c('Leaf defense vs. investment',
               'Leaf ressource capture vs.investment',
               'Leaf defense vs. investment',
               'Leaf ressource acquisiton vs. defense',
               'Stem transport, structure and defense')
  )
kable(table,  label="traits", row.names = F)
```

@@

```{r 2 CV}
PFT_ind <- aggregate(PFT, by = list(PFT$Tree), FUN = mean, na.rm = T)
row.names(PFT_ind) <- PFT_ind$Group.1
SpT <- unique(PFT[1:2])
Sp <- SpT$SP
names(Sp) <- SpT$Tree
SpCodeT <- unique(PFT[2:3])
SpCode <- SpCodeT$Sp_Code
names(SpCode) <- SpCodeT$SP
CET <- unique(PFT[c(1,4)])
CE <- CET$CE
names(CE) <- CET$Tree
PFT_ind$SP <- Sp[row.names(PFT_ind)]
PFT_ind$Sp_Code <- SpCode[as.character(PFT_ind$SP)]
PFT_ind$CE <- CE[row.names(PFT_ind)]
PFT_ind <- PFT_ind[-c(1:2)]
cv <- data.frame(row.names = c('Thick', 'LA', 'LDMC', 'SLA', 'WD'))
cv$Individual <- unlist(lapply(
  as.list(row.names(cv)),
  function(y){mean(aggregate(PFT[which(names(PFT) ==  y)], by = list(PFT$Tree), function(x){sd(x, na.rm = T) / mean(x, na.rm = T)})[,2], na.rm = T)}))
cv$Species <- unlist(lapply(
  as.list(row.names(cv)),
  function(y){mean(aggregate(PFT_ind[which(names(PFT_ind) ==  y)], by = list(PFT_ind$SP), function(x){sd(x, na.rm = T) / mean(x, na.rm = T)})[,2], na.rm = T)}))
cv$Community <- unlist(
  lapply(as.list(row.names(cv)),
         function(y){sd(PFT[,which(names(PFT) ==  y)], na.rm = T) /
             mean(PFT[,which(names(PFT) ==  y)], na.rm = T)}))
cv <- round(cv, 2)
cv['WD','Individual'] <- ''
kable(cv)
```

@@

```{r 3 PCA PFT}
library(vegan)
pca <- princomp(~ LA + WD + Thick + LDMC + SLA, data = PFT_ind, cor = T)
lambda <- pca$sdev * sqrt(nrow(pca$scores))
pca.var = apply(pca$score, 2, var)
PCAscores <- as.data.frame(pca$scores)[1:2] # Adding PCA axis in PFT values
names(PCAscores) <- c('WES', 'LES')
PCAscores$ID <- row.names(PCAscores)
plot(t(t(pca$scores)/lambda),pch=16, col = 'lightgrey',
     # xlab = paste('Axe 1 -', round(lambda[1]/sum(lambda)*100, 2), '%'),
     # ylab = paste('Axe 2 -', round(lambda[2]/sum(lambda)*100, 2), '%'))
     xlab = paste('Axe 1 -', round(pca.var[1]/sum(pca.var)*100, 2), '%'),
     ylab = paste('Axe 2 -', round(pca.var[2]/sum(pca.var)*100, 2), '%'))
# as.factor(Species[PFT_ind[row.names(pca$scores),2],6])
par(new=T)
Rot <- t(t(pca$loadings)*lambda)
XLIM <- c(-max(abs(Rot[,1])),max(abs(Rot[,1])))
XLIM <- XLIM+(XLIM*0.2)
plot(Rot,col=4,axes=FALSE,xlim=XLIM,ylim=XLIM,pch="",xlab = "", ylab = "")
lines(XLIM, c(0,0), lty = 2, col = 'firebrick', lwd = 2)
lines(c(0,0), XLIM, lty = 2, col = 'green', lwd = 2)
legend('topright', c('Leaf economic spectrum', 'Wood economic spectrum'),
       text.col = c("green", "firebrick"))

@@

```{r 4 env-com}
library(raster)
Env <- genEnv()
Env$Rivers <- readAll(Env$Rivers)
Trees <- genTrees()
Species <- genSpecies()
PFT_sp <- aggregate(PFT_ind, by = list(PFT_ind$SP), FUN = mean, na.rm = T)
row.names(PFT_sp) <- PFT_sp$Group.1
PFT_sp$Sp_Code <- SpCode[row.names(PFT_sp)]
PFT_sp <- PFT_sp[-c(1:2,4)]
Trees$SpCode <- tolower(Trees$SpCode)
Trees$ID <- row.names(Trees)
Trees <- merge(Trees, PFT_sp, by.x = 'SpCode', by.y = 'Sp_Code')
row.names(Trees) <- Trees$ID
Trees <- Trees[-which(names(Trees) == 'ID')]
com <- quadrats()
Trees$com <- NA
XY <- data.frame(X = Trees$X, Y = Trees$Y)
XY <- XY[-which(is.na(XY$X)),] # Removing NA
coordinates(XY) <- ~ X + Y
proj4string(XY) <- crs(com)
Trees$com[which(!is.na(Trees$X))] <- (XY %over% com)[,1]
Trees$com[which(!is.na(Trees$com))] <- paste0('C', Trees$com[which(!is.na(Trees$com))])
com@data$id <- paste0('C', com@data$id)
com@data$DEM <- extract(Env$DEM, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
Deriv <- DEMderiv(raster(com, 'DEM'), c('slope', 'curvature', 'plancurvature', 'profcurvature', 'aspect'))
com@data <- cbind(com@data, extract(Deriv, as(com, "SpatialPolygons"), df = T, fun = mean)[-1])
com@data$Wetness <- extract(DEMderiv(Env$DEM, 'wetness'), as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$SouthWesterness <- sin((com@data$Aspect + 225) / 180 * pi)
com@data$NorthEasterness <- cos((com@data$Aspect + 225) / 180 * pi)
com@data$Rivers <- extract(Env$Rivers, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$topo <- as.factor(with(com@data, topo_classes(Slope, Curvature, Aspect)))
com@data$Canopy <- extract(Env$Canopy, as(com, "SpatialPolygons"), df = T, fun = mean)[,2]
com@data$BA <- aggregate(Trees$`2013_girth`, list(Trees$com), function(x){sum(x, na.rm = T)/2*pi})[,2]
com@data$StD <- aggregate(rep(1, length(Trees$SpCode)), list(Trees$com), sum)[,2]
rm(Deriv, XY)
```

@@

To study biotic factors we choose 3 available variables: canopy height (canopy), basal area (BA), and stem density (StD) *(we may add tree height and growth in the future)*. Canopy height is issued from a digital canopy model built with a LIDAR campaign on Uppangala plots with a resolution $1*1 m^2$. Both basal area and stem density are calculated with 2013 Uppangala plots girth inventory only on stem with a diameter above 10 centimeters at breast height ($DBH>10$). Consequently basal area end stem density are highly correlated (Spearman's $rho = 0.84$, $p-value < 0.001$). So we removed stem density to avoid multicolinearity, and we kept canopy height that wasn't correlated to basal area (Spearman's $rho = -0.09$, $p-value > 0.1$).

For each species the potential strata was defined. Because species were defined in several strata, we kept the lowest strata as the species strata. We tested the strata effect on each trait with an anova combined to a tukey test, but no trait show a significant variation (all F values weren't significanlty different from 0, $p-values>0.1$).

```{r strata effect}
AOV <- list(SLA = aov(PFT_sp$SLA ~ Species[PFT_sp$Sp_Code,]$Strata), 
            WD = aov(PFT_sp$WD ~ Species[PFT_sp$Sp_Code,]$Strata),
            LDMC = aov(PFT_sp$LDMC ~ Species[PFT_sp$Sp_Code,]$Strata),
            LA = aov(PFT_sp$LA ~ Species[PFT_sp$Sp_Code,]$Strata),
            Thick = aov(PFT_sp$Thick ~ Species[PFT_sp$Sp_Code,]$Strata))
# lapply(AOV, summary)
# lapply(AOV, TukeyHSD)
```

@@

```{r 5 Com}
com <- CWT(Trees, com)
```

Among the whole 10 hectares plot from Uppangala, 240 spatial communities where defined with $20*20 m^2$ quadrats. For all those community, community weighted means and variances [CWM and CWV, sensu @violle_let_2007] were computed for each functional traits and from the species scores obtained on the 2 first PCA axes representing the leaf and the wood economic spectrum (WES, LES). Weighted mean and variances were realized by abundance, presence-absence, or basal area.

Spatial variograms of community weighted means and variances didn't show a clear plateau (see figure 4) but suggest a spatial organization at a scale of $70 m$.

```{r Variograms}
library(gstat)
dist <- variogram(WD.cwm~1, as(com$abundance, 'SpatialPointsDataFrame'))$dist
par(mfrow = c(2,1))
for(i in 1:2){
  traits <- paste0(c('WD', 'Thick', 'LA', 'LDMC', 'SLA', 'WES', 'LES'), c('.cwm','.cwv')[i])
  plot(c(1:length(dist)) ~ dist, xlim = c(0, max(dist)), ylim = c(0.5,1), pch = '', main = c('(a)', '(b)')[i], xlab = 'distance (m)', ylab = 'relative variance (%)')
  for(j in 1:length(com)){
    var <- data.frame(lapply(lapply(as.list(traits), function(x){variogram(formula(paste(x, '~ 1')), as(com[[j]], 'SpatialPointsDataFrame'))}), function(x){x[3]}))
    var <- data.frame(apply(var, 2, function(x){x/max(x)}))
    names(var) <- c('WD', 'Thick', 'LA', 'LDMC', 'SLA', 'WES', 'LES')
    colramp <- rainbow(length(var))
    for(i in 1:length(var)){
      lines(var[,i] ~ dist, col = colramp[i], lty = j + 1)
    }
  }
legend('bottomright', names(var), fill = colramp, cex = 0.7)
legend('bottomleft', names(com), lty = c(2:4), cex = 0.7)
}
```

Figure 4: Trait variograms for all communities.*Variograms were initiated on* $20*20 m^2$ *communities for all traits with both community weighted means* **(a)** *and variances* **(b)** *(by abundance, presence-absence and basal area). Some variaograms seem to reach a plateau around* $70 m$.

In order to see if environmental biotic factors were influencing functionnal traits or if environmental biotic factors were the results of functionnal traits we used structural equation modelling on community weighted means by abundance. Three latent variables were defined: *Abiotic environment* defined with our 4 selected topographical variables, *Biotic environment* defined with our 2 selected biotic variables, and *Functionnal traits* defined with our 5 functionnal traits. Two models were compared, a first one with the functionnal traits explained by both abiotic and biotic factors (**m1** *Traits ~ Abiotic + Biotic*), and a second one with the biotic factors explained by the functionnal traits and abiotic factors (**m2** *Biotic ~ Traits + Abiotic*). Retained model was the first one (**m1**) with a significantly lower Aikaike's Criterion Indice on Chi Square Difference test ($AIC = 13195$ against $AIC = 13238$ with $p.value < 0.001$)

```{r SEM}
library(lavaan)
m1 <- '
Abiotic =~ Slope+Curvature+Wetness+SouthWesterness
Traits =~ WD.cwm+Thick.cwm+LA.cwm+LDMC.cwm+SLA.cwm
Biotic =~ BA+Canopy
Traits ~ Abiotic+Biotic
'
m2 <- '
Abiotic =~ Slope+Curvature+Wetness+SouthWesterness
Traits =~ WD.cwm+Thick.cwm+LA.cwm+LDMC.cwm+SLA.cwm
Biotic =~ BA+Canopy
Biotic ~ Traits +Abiotic
'
fit1 <- sem(m1, data = com$abundance@data)
fit2 <- sem(m2, data = com$abundance@data)
semPlot::semPaths(fit1,what='path',whatLabels='std',intercept=FALSE,residuals=FALSE,thresholds=FALSE,nCharNodes=0,nCharEdges=0,groups='manlat',pastel=TRUE,title=TRUE,curveAdjacent=TRUE,ask=FALSE)
```

Figure 5: Path diagram for structural equation modelling. *Path diagram of the structural equation model* **m1** *Traits ~ Abiotic + Biotic*.

Links between both community weighted means and variances with their environment, for both abiotic and biotic factors, were studied by linear regressions on continuous variables. Linear regression models were chosen by a step mesure (in both direction) on a first regression with selected variables (slope, curvature, wetness and southwesterness for abiotic factors, and canopy height and basal area for biotic one). A null model was built by randomizing traits among species (permuting raw in `Trait*Species` matrix). Deviance of models selected by the step mesure was calculated 999 times and compare to observed deviance. Results for abundance, presence-absence and basal area community weighted mean for all traits are presented in tables 4 and 5 (respectively for community weighted means and variances).

An assessment of spatial autocorrelation (SAC) of the community weighted means was conducted using Moran's I statistic. Most of community weighed means, especially in abundance, were significantly positively spatially autocorrelated (see table 3). But half of presence-absence and basal area community weighted means spatial autocrorrelation didn't significantly differ from 0 and the other were all low (Moran's I from 0.01 to 0.03). Specific leaf area and wood density are slightly spatially correlated suggesting a spatial filter due to environment or local history pressure (as dispersion limitation). Leaf dry mater content and more generally the wood economic spectrum possess a stochastic spatial distribution implying that they're not shaped by factors with a spatial organization. Finally the leaf economic spectrum, thickness and area are spatially organized for abundant species and voluminous individual but not in presence absence, suggesting that the filter is linked to abundant species pressure as dispersion filter.

```{r SAC}
library(ape)
traits <- c('WD', 'Thick', 'LA', 'LDMC', 'SLA', 'WES', 'LES')
traits <- c(paste0(traits, '.cwm'), paste0(traits, '.cwv'))
weights <- 1 / as.matrix(dist(coordinates(com$abundance)))
diag(weights) <- 0
SAC <- lapply(com, function(x){lapply(as.list(x@data[traits]), Moran.I, weights)})
SAC <- lapply(SAC, function(x){do.call(rbind.data.frame, x)})
SAC <- data.frame(lapply(SAC, function(x){paste(round(x$observed, 2), unlist(lapply(as.list(x$p.value), stars)))}))
names(SAC) <- c('Abundance', 'Presence-absence', 'Basal area')
SAC$Trait <- c('LES', 'SLA', 'WES', 'WD', 'LDMC', 'Thick', 'LA')
SAC$Type <- ''
SAC$Type[c(1,8)] <- c('CWM', 'CWV')
SAC <- SAC[c('Type', 'Trait', 'Abundance', 'Presence-absence', 'Basal area')]
kable(SAC, row.names = F)
```

Table 3: Spatial autocorrelation measured by Moran's I for community weighted means and variances. *Spatial autocorrelation for each trait community weighted means and variances (by abundance, presence-absence and basal area), see table 1 for abbreviations.*

```{r 11 lm & null models}
reg.abiot <- '~ Slope + Curvature + Wetness + SouthWesterness'
reg.biot <- '~ Canopy + BA'
# LM to simplify with loop or lapply
LM <- lapply(com, function(x) {with(x@data, list(LES.abiot.cwm = lm(formula(paste('LES.cwm', reg.abiot))), # Abiotic CWM
                                                 logSLA.abiot.cwm = lm(formula(paste('log(SLA.cwm)', reg.abiot))), 
                                                 WES.abiot.cwm = lm(formula(paste('WES.cwm', reg.abiot))),
                                                 WD.abiot.cwm = lm(formula(paste('WD.cwm', reg.abiot))),
                                                 LDMC.abiot.cwm = lm(formula(paste('LDMC.cwm', reg.abiot))),
                                                 logLA.abiot.cwm = lm(formula(paste('log(LA.cwm)', reg.abiot))),
                                                 Thick.abiot.cwm = lm(formula(paste('Thick.cwm', reg.abiot))),
                                                 LES.abiot.cwv = lm(formula(paste('LES.cwv', reg.abiot))), # Abiotic CWV
                                                 logSLA.abiot.cwv = lm(formula(paste('log(SLA.cwv)', reg.abiot))), 
                                                 WES.abiot.cwv = lm(formula(paste('WES.cwv', reg.abiot))),
                                                 WD.abiot.cwv = lm(formula(paste('WD.cwv', reg.abiot))),
                                                 LDMC.abiot.cwv = lm(formula(paste('LDMC.cwv', reg.abiot))),
                                                 logLA.abiot.cwv = lm(formula(paste('log(LA.cwv)', reg.abiot))),
                                                 Thick.abiot.cwv = lm(formula(paste('Thick.cwv', reg.abiot))),
                                                 LES.biot.cwm = lm(formula(paste('LES.cwm', reg.biot))), # Biotic CWM
                                                 logSLA.biot.cwm = lm(formula(paste('log(SLA.cwm)', reg.biot))),
                                                 WES.biot.cwm = lm(formula(paste('WES.cwm', reg.biot))),
                                                 WD.biot.cwm = lm(formula(paste('WD.cwm', reg.biot))),
                                                 LDMC.biot.cwm = lm(formula(paste('LDMC.cwm', reg.biot))),
                                                 logLA.biot.cwm = lm(formula(paste('log(LA.cwm)', reg.biot))),
                                                 Thick.biot.cwm = lm(formula(paste('Thick.cwm', reg.biot)))
                                                 ))})
LM <- lapply(LM, function(y){lapply(y, function(x){step(x, direction = 'both', trace = 0)})})
model <- null_model(com, Trees, PFT_sp, lapply(LM, function(y){lapply(y, function(x){x$call})}), SpCode, n = 999, verbose = F, time = T)
```

@@

```{r CWM abiotic results}
kable(regression_table(LM, model, reg.abiot, select = '.abiot.cwm'))
```

Table 4: Step mesure on abiotic factors and null models summary for traits community weighted means. *Linear model results for each trait community weighted means (by abundance, presence-absence and basal area) after a variable selection by step mesure in both direction with slope, curvature, wetness, and southwesterness (SW). Model deviance was compared to 999 null models deviance in which traits were randomized among species.*

@@

An assessment of spatial autocorrelation (SAC) of model residuals confirm that most of the spatial structure of our traits is not explained by our models because the spatial organization stay the same as the one found on the community weighted mean shown in table 3.

```{r CWV biotic results}
kable(regression_table(LM, model, reg.abiot, select = '.abiot.cwv'))
```

Table X: Abiotic CWV

## Individual filters
